Dismal Code: Studying the Evolution of Security Bugs

### 2013-04-07 docs/evol.tex
started the discussion section

\author{Dimitris Mitropoulos \and Vassilios Karakoidas \and Georgios Gousios \and Panos Louridas \and Diomidis Spinellis }
\institute{ Department of Management Science and Technology\\
			Athens University of Economics and Business\\
      		\email{\{dimitro, bkarak, louridas, dds\}@aueb.gr}\\
			Software Engineering Research Group\\
			Delft University of Technology\\
			\email{G.Gousios@tudelft.nl}}

@@

A security bug is a programming error that introduces a potentially
exploitable weakness into a computer system~\cite{SSL12}\cite{TJBD11}. This weakness could lead to a
security breach with unfortunate consequences in different layers, like databases,
native code, applications, libraries and others. Despite the significant
effort to detect and eliminate such bugs~\cite{SZ12}, little attention has been paid to
study them in relation to software evolution~\cite{L96}\cite{LRWPT97}\cite{IB06}\cite{RGMA06}.
In this paper we present how we used a large software ecosystem to analyse
how evolving software packages are related to the different types of security
vulnerabilities.

@@

The motivation behind our work was to validate if programmers tend to care for
such high-risk bugs when they release a new version of their software. In
addition, we wanted to investigate other critical features associated with such
vulnerabilities like the persistence of a bug. In essence, to check if critical bugs stay
unresolved for a long time. Also, we wanted to elaborate more on the relation of security
bugs not only with performance bugs (as Zaman et al.~\cite{ZAH11} have already
did) but also with other bug categories. In the same manner, we tried to check
the relation between a version's size and its' security bugs knowing the
that research has been contradictory on the the
issue~\cite{BP84}\cite{SYTP85}\cite{NBZ06}\cite{GKMS00}.
Finally, we examined the Maven ecosystem as a whole from a security
perspective. Its structure, gave us the opportunity to see if a project version that is a dependency to
a large number of others contains a high rate of security bugs.
In this case we can strengthen the relationship between Linus' Law and
software security as Meneely et al. have already indicated~\cite{MW10}.

@@

Figure~\ref{fig:bug-per} shows how software bugs are distributed among the
repository. Together with the {\it Bad Practice} bugs and the {\it Style} bugs,
security bugs (the sum of the {\it Security} and {\it Malicious Code}
categories) are the most popular in the repository ($\geq 22\%$).
This could be an indication that programmers write code that implements the
required functionality without considering its many security aspects.
% An interesting observation is that 4,908 projects did not
% have any bugs at all. These cases involved small artifacts with a minimum number of classes.
% This goes along with the relationship between the size of an artifact and it's
% bugs that is presented in the upcoming section.

@@

As we mentioned earlier, during the experiment we managed to retrieve the
dependencies of every version. Based on this information we created a graph
that represented the snapshot of the Maven repository. The
nodes of the graph represented the versions and the vertices their dependencies.
We have to mention here that the graph was not representative. For instance, if
a dependency was pointing only to a project (and not to a specific version), we chose to
select the latest version found on the repository. Also, this graph is not
complete. This is because there were missing versions.
% Practically, the corresponding {\it pom}, or {\it war} files existed on the
% repository, but the {\sc jar} file did not.
From the 56,5680 vertices, 191,433
belonged to the first case while 16,4234 belonged to the second.
The graph contained 80,354 nodes. Obviously, the number does not correspond to
the number of the total versions (see Section~\ref{sec:data}). This is because
some versions did not contain any information about their dependencies so they
are not represented in the graph. After creating the graph, we ran the PageRank
algorithm~\cite{BP98} on it and retrieved all PageRanks for every node. Then we
examined the security bugs of the fifty most popular nodes based on their PageRank.
Thirty three of them contained bugs coming from the {\it Security Low} subcategory,
while two of them contained severe bugs. Twenty five of them were latest
versions at the time.
% what can we say about this?

@@

>>>>

\author{Dimitris Mitropoulos \and Vassilios Karakoidas \and Panos Louridas \and Georgios Gousios \and Diomidis Spinellis }
\institute{ Department of Management Science and Technology\\
			Athens University of Economics and Business\\
      		\email{\{dimitro, bkarak, louridas, dds\}@aueb.gr}\\
			Software Engineering Research Group\\
			Delft University of Technology\\
			\email{G.Gousios@tudelft.nl}}

@@

A security bug is a programming error that introduces a potentially
exploitable weakness into a computer system~\cite{SSL12}\cite{TJBD11}\cite{M06}. This weakness could lead to a
security breach with unfortunate consequences in different layers, like databases,
native code, applications, libraries and others. Despite the significant
effort to detect and eliminate such bugs~\cite{SZ12}, little attention has been paid to
study them in relation to software evolution~\cite{L96}\cite{LRWPT97}\cite{IB06}\cite{RGMA06}.
In this paper we present how we used a large software ecosystem to analyse
how evolving software packages are related to the different types of security
vulnerabilities.

@@

The motivation behind our work was to validate if programmers tend to care for
such high-risk bugs when they release a new version of their software. In
addition, we wanted to investigate other critical features associated with such
vulnerabilities like the persistence of a bug. In essence, to check if critical bugs stay
unresolved for a long time. Also, we wanted to elaborate more on the relation of security
bugs not only with performance bugs (as Zaman et al.~\cite{ZAH11} have already
did) but also with other bug categories. In the same manner, we tried to check
the relation between a version's size and its' security bugs knowing the
that research has been contradictory on the the
issue~\cite{BP84}\cite{SYTP85}\cite{NBZ06}\cite{GKMS00}.
Finally, we examined the Maven ecosystem as a whole from a security
perspective. Its structure, gave us the opportunity to see if a project version that is a dependency to
a large number of others contains a low rate of security bugs~\cite{MW10}.
%In this case we can strengthen the relationship between Linus' Law and
%software security as Meneely et al. have already indicated~\cite{MW10}.

@@

Figure~\ref{fig:bug-per} shows how software bugs are distributed among the
repository. Together with the {\it Bad Practice} bugs and the {\it Style} bugs,
security bugs (the sum of the {\it Security} and {\it Malicious Code}
categories) are the most popular in the repository ($\geq 22\%$).
% An interesting observation is that 4,908 projects did not
% have any bugs at all. These cases involved small artifacts with a minimum number of classes.
% This goes along with the relationship between the size of an artifact and it's
% bugs that is presented in the upcoming section.

@@

As we mentioned earlier, during the experiment we managed to retrieve the
dependencies of every version. Based on this information we created a graph
that represented the snapshot of the Maven repository. The
nodes of the graph represented the versions and the vertices their dependencies.
We have to mention here that the graph was not representative. For instance, if
a dependency was pointing only to a project (and not to a specific version), we chose to
select the latest version found on the repository. Also, this graph is not
complete. This is because there were missing versions.
% Practically, the corresponding {\it pom}, or {\it war} files existed on the
% repository, but the {\sc jar} file did not.
From the 565,680 vertices, 191,433
belonged to the first case while 164,234 belonged to the second.
The graph contained 80,354 nodes. Obviously, the number does not correspond to
the number of the total versions (see Section~\ref{sec:data}). This is because
some versions did not contain any information about their dependencies so they
are not represented in the graph. After creating the graph, we ran the PageRank
algorithm~\cite{BP98} on it and retrieved all PageRanks for every node. Then we
examined the security bugs of the fifty most popular nodes based on their PageRank.
Thirty three of them contained bugs coming from the {\it Security Low} subcategory,
while two of them contained severe bugs. Twenty five of them were latest
versions at the time.
% what can we say about this?

@@

By examining Maven as a whole we observed that security bugs and the
bugs coming from the {\it Bad Practice} category are the most popular types of
bugs in the ecosystem. This could be a strong indication that programmers write code
that implements the required functionality without considering its many
security aspects; an issue that has already been reported in
literature~\cite{SH09}. In addition, contrary to the Linus' Law, from the top fifty versions
(based on their PageRank), twenty five of them were the latest versions at the
time and at the same time contained security bugs.
This also highlights the {\it domino effect}~\cite{TH04}. No matter how well a programmer
secures a software component it won't matter if he or she is using vulnerable
code fragments from another library. Finally, 5,355 versions ($\approx 4,64\% $), contain at
least one severe security bug. Given the fact that other projects include these
versions as their dependencies, they are automatically rendered vulnerable if
they use the code fragments that include the defects.
